\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-08-04 paper/paper.tex
Update paper

\begin{equation}
\label{eq:model}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
\end{equation}

where \(\beta_{0,j}\) is the diameter-independent growth rate of species
\(j\); \(\text{dbh}_{ij}\) is the DBH of the focal tree at the earlier
census and \(\beta_{\text{dbh},j}\) the slope of that species's
diameter-growth relationship; \(\text{BA}_{ijk}\) is the sum of the
basal area of all trees of competitor species \(k\) and \(\lambda_{jk}\)
quantifies the corresponding change in growth for individuals of group
\(j\) from these competitors; and \(\epsilon_{ij}\) is a random error
term distributed \(\text{Normal}(0, \sigma^2)\).
\citet{allen_permutation_2020} estimate all parameters via Bayesian
linear regression, while exploiting Normal/Inverse Gamma conjugacy to
derive closed-form solutions to all posterior distributions\footnote{See
  S1 Appendix of \citet{allen_permutation_2020}, available at
  \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
closed-form solutions are not as computationally expensive as
approximations from Markov Chain Monte Carlo algorithms.

@@

The package is designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like all \texttt{tidyverse} packages,
\texttt{forestecology} has verb-named functions that can be modularly
composed using the pipe \texttt{\%\textgreater{}\%} operator to
sequentially complete all necessary analysis steps
\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most
functions use the same ``simple features for R'' data structures from
the \texttt{sf} package, a package for standardized and
\texttt{tidyverse}-friendly wrangling and visualizing of spatial data
\citep{pebesma_simple_2018}.

Currently the package only implements the Bayesian linear regression
model detailed in Equation \ref{eq:model}. As we demonstrate in Section
\ref{model-fit-predict} however, the fitting of this model is
self-contained in a single function \texttt{comp\_bayes\_lm()} which
returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
has generic methods implemented to print, make predictions, and plot all
results. Therefore the package can be modularly extended to fit other
models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
and have equivalent generic methods implemented.

@@

We load both 2008 and 2014 SCBI census \texttt{.csv} files as they
existed on GitHub on 2020/11/20 and perform minor data wrangling
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. We then
only consider a 9 ha subsection of the 25.6 ha of the site to speed up
computation for this example: \texttt{gx} from 0--300 instead of 0--400
and \texttt{gy} from 300--600 instead of 0--640.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
\CommentTok{## Simple feature collection with 7954 features and 5 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 6}
\CommentTok{##   stemID sp     dbh1  dbh2 growth   geometry}
\CommentTok{##    <dbl> <fct> <dbl> <dbl>  <dbl>    <POINT>}
\CommentTok{## 1      4 nysy  13.6   14.2  0.103 (14.2 428)}
\CommentTok{## 2      5 havi   8.8    9.6  0.150  (9.4 436)}
\CommentTok{## 3      6 havi   3.25   4    0.140  (1.3 434)}
\CommentTok{## 4     77 qual  65.2   66    0.141 (34.7 307)}
\CommentTok{## 5     79 tiam  47.7   46.8 -0.161   (40 381)}
\CommentTok{## # ... with 7,949 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{,y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\},}\StringTok{')'}\NormalTok{)) )}
\end{Highlighting}
\end{Shaded}

@@

@@

\CommentTok{# Manually construct study region boundary}
\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}

@@

\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}

\CommentTok{# Associate each observation to a fold}
\NormalTok{spatial_block_scbi <-}\StringTok{ }
\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }
               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
\CommentTok{## # A tibble: 6,296 x 5}
\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
\CommentTok{## # A tibble: 20 x 6}
\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area}
\CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>             <dbl>}
\CommentTok{## 1        4  13.6    1836  7.48 tiam            0.0176 }
\CommentTok{## 2        4  13.6    1847  2.81 nysy            0.00332}
\CommentTok{## 3        4  13.6    1848  1.62 nysy            0.00396}
\CommentTok{## 4        4  13.6    1849  2.62 nysy            0.00535}
\CommentTok{## 5        4  13.6    1850  2.98 havi            0.00472}
\CommentTok{## # ... with 15 more rows}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }
                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }
                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }
                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}

\CommentTok{# Combine plots using the patchwork package}
\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }
                                 \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{forestecology} package provides an accessible way to fit and
test models of neighborhood competition. The package follows the
\texttt{tidy} data design principles, leverages the \texttt{sf} package
for spatial data, and S3 open-oriented model implementation structure.
We hope that the package will increase the use of neighborhood
competition models to better understand what structures plant
competition.

@@

\CommentTok{# Run all simulations}
\CommentTok{# 0. Setup simulation for this species type ----}
\CommentTok{# Start clock}
\KeywordTok{tic}\NormalTok{()}

>>>>

\begin{equation}
\label{eq:model}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
\end{equation}

where \(\beta_{0,j}\) is the diameter-independent growth rate of species
\(j\); \(\text{dbh}_{ij}\) is the DBH of the focal tree at the earlier
census and \(\beta_{\text{dbh},j}\) the slope of that species's
diameter-growth relationship; \(x^{\text{comp}}_{ijk}\) is the sum of
some numerical explanatory variable of all trees of competitor species
\(k\) and \(\lambda_{jk}\) quantifies the corresponding change in growth
for individuals of group \(j\) from these competitors; and
\(\epsilon_{ij}\) is a random error term distributed
\(\text{Normal}(0, \sigma^2)\).

\citet{allen_permutation_2020} use the sum of the basal area of all
trees of competitor species \(k\) as \(x^{\text{comp}}_{ijk}\).
Furthermore, they estimate all parameters via Bayesian linear
regression, while exploiting Normal/Inverse Gamma conjugacy to derive
closed-form solutions to all posterior distributions\footnote{See S1
  Appendix of \citet{allen_permutation_2020}, available at
  \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
closed-form solutions are not as computationally expensive as
approximations from Markov Chain Monte Carlo algorithms.

@@

The package is designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like all \texttt{tidyverse} packages,
\texttt{forestecology} has verb-named functions that can be modularly
composed using the pipe \texttt{\%\textgreater{}\%} operator to
sequentially complete all necessary analysis steps
\citep{bache_pipe_2020}.

Furthermore, the inputs and outputs of most functions use the same
``simple features for R'' data structures for spatial data from the
\texttt{sf} package \citep{pebesma_simple_2018}. Previously \texttt{sp}
package classes were commonly used for storing spatial data and
interfacing with geospatial libraries \citep{bivand_sp_2013}; the
\texttt{sf} package aims to improve on the \texttt{sp} package by:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Using simple feature access as the base standard for representing and
  encoding spatial data, rather than shapefiles \citep{herring_2011}.
\item
  Leveraging improvements in external libraries for reading and writing
  spatial data (GDAL) and for geometrical operations (GEOS)
  \citep[\citet{geos_2017}]{warmerdam_geospatial_2008}.
\item
  Integrating closely with the popular \texttt{tidyverse} suite of
  packages for data science \citep{wickham_welcome_2019}.
\end{enumerate}

By using the \texttt{sf} package classes to represent spatial data
rather than the \texttt{sp} package, the implementation and use of the
\texttt{forestecology} package's spatial algorithms was greatly
simplified.

Lastly, currently the package only implements the Bayesian linear
regression model detailed in Equation \ref{eq:model}. As we demonstrate
in Section \ref{model-fit-predict} however, the fitting of this model is
self-contained in a single function \texttt{comp\_bayes\_lm()} which
returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
has generic methods implemented to print, make predictions, and plot all
results. Therefore the package can be modularly extended to fit other
models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
and have equivalent generic methods implemented.

@@

We load both 2008 and 2014 SCBI census \texttt{.csv} files as they
existed on GitHub on 2021/08/02 and perform minor data wrangling
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. We then
only consider a 9 ha subsection of the 25.6 ha of the site to speed up
computation for this example: \texttt{gx} from 0--300 instead of 0--400
and \texttt{gy} from 300--600 instead of 0--640.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
\CommentTok{## Simple feature collection with 7954 features and 5 fields}
\CommentTok{## Geometry type: POINT}
\CommentTok{## Dimension:     XY}
\CommentTok{## Bounding box:  xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
\CommentTok{## CRS:           NA}
\CommentTok{## # A tibble: 7,954 x 6}
\CommentTok{##   stemID sp     dbh1  dbh2 growth   geometry}
\CommentTok{##    <dbl> <fct> <dbl> <dbl>  <dbl>    <POINT>}
\CommentTok{## 1      4 nysy  13.6   14.2  0.103 (14.2 428)}
\CommentTok{## 2      5 havi   8.8    9.6  0.150  (9.4 436)}
\CommentTok{## 3      6 havi   3.25   4    0.140  (1.3 434)}
\CommentTok{## 4     77 qual  65.2   66    0.141 (34.7 307)}
\CommentTok{## 5     79 tiam  47.7   46.8 -0.161   (40 381)}
\CommentTok{## # ... with 7,949 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}
    \DecValTok{1}\NormalTok{)) }\OperatorTok{+}\StringTok{ }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}
    \DecValTok{-1}
\NormalTok{\}, }\StringTok{")"}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

@@

We also load species information as it existed on GitHub on 2021/08/02,
which includes family, genus, and species information; as well as
classifications of the canopy position (canopy, canopy emergent,
understory, shrub layer) and drought tolerance (intolerant, resistant)
of the species.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sp_info <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"SCBI_ForestGEO_sp_ecology.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, drought_tolerance) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{()}
\NormalTok{sp_info}
\CommentTok{## # A tibble: 65 x 6}
\CommentTok{##   sp    family        genus       species     canopy_position drought_toleran~}
\CommentTok{##   <chr> <chr>         <chr>       <chr>       <chr>           <chr>           }
\CommentTok{## 1 acne  Sapindaceae   Acer        negundo     understory      drought-resista~}
\CommentTok{## 2 acpl  Sapindaceae   Acer        platanoides canopy          drought-resista~}
\CommentTok{## 3 acru  Sapindaceae   Acer        rubrum      canopy          drought-resista~}
\CommentTok{## 4 aial  Simaroubaceae Ailanthus   altissima   canopy          drought-resista~}
\CommentTok{## 5 amar  Rosaceae      Amelanchier arborea     understory      drought-resista~}
\CommentTok{## # ... with 60 more rows}
\end{Highlighting}
\end{Shaded}

We join this species information to our \texttt{growth\_scbi} data
frame.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
\end{Highlighting}
\end{Shaded}

Furthermore, we compute two potential competitor explanatory variables
\(x^{\text{comp}}_{ijk}\) from Equation \ref{eq:model}. First, the basal
area of each tree as a function of its DBH in the earlier census.
Second, the above ground biomass as estimated by allometric equations
encoded in the \texttt{get\_biomass()} function from the \texttt{allodb}
package \citep{allodb_package}; this function has DBH, species, and
geographic coordinates as arguments.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Install development version of allodb using:}
\CommentTok{# remotes::install_github("forestgeo/allodb")}
\KeywordTok{library}\NormalTok{(allodb)}
\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \CommentTok{# Compute basal area:}
    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
    \CommentTok{# Compute above ground biomass:}
    \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
      \DataTypeTok{dbh =}\NormalTok{ dbh1,}
      \DataTypeTok{genus =}\NormalTok{ genus,}
      \DataTypeTok{species =}\NormalTok{ species,}
      \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
\NormalTok{    )}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{# Manually construct study region boundary}
\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}

@@

\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{factor}\NormalTok{())}

\CommentTok{# Associate each observation to a fold}
\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{factor}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{comp_x_var =}\NormalTok{ basal_area)}

\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
\CommentTok{## # A tibble: 6,296 x 5}
\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
\CommentTok{## # A tibble: 20 x 6}
\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_x_var}
\CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>        <dbl>}
\CommentTok{## 1        4  13.6    1836  7.48 tiam       0.0176 }
\CommentTok{## 2        4  13.6    1847  2.81 nysy       0.00332}
\CommentTok{## 3        4  13.6    1848  1.62 nysy       0.00396}
\CommentTok{## 4        4  13.6    1849  2.62 nysy       0.00535}
\CommentTok{## 5        4  13.6    1850  2.98 havi       0.00472}
\CommentTok{## # ... with 15 more rows}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}

\CommentTok{# Combine plots using the patchwork package}
\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2)}\OperatorTok{/}\NormalTok{plot3}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{forestecology} package provides an accessible way to fit and
test models of neighborhood competition. The package follows the
\texttt{tidy} data design principles, leverages the \texttt{sf} package
for spatial data, and S3 open-oriented model implementation structure
\citep{pebesma_simple_2018}. We hope that the package will increase the
use of neighborhood competition models to better understand what
structures plant competition.

@@

\CommentTok{# Run all simulations 0. Setup simulation for this species type ----}
\CommentTok{# Start clock}
\KeywordTok{tic}\NormalTok{()}
